AXIELL piskoviste

### 2023-10-10 _GEN_axi_Rhacks.Rmd
Long time no commit 19/7 + MUZTE init
po migraci MUZTE, uprava gen CSV
CES MUZEUMDC - reorg.

### Read txt

read.table

```{r READ TABLE}
df <- read.table("D:/VURV/R_hacks/data/simple_table_4read.txt", header = TRUE, sep = ";")
```

### Read csv

@@

```{r READ CSV}
df <- read.csv("D:/VURV/R_hacks/data/simple_table_4read.csv", header = TRUE, sep = ";")
```

### Read excel

```{r READ EXCEL}

library("readxl")
df <- read_excel("D:/VURV/R_hacks/data/simple_table_4read.xlsx" , sheet = 2, header = TRUE)
```

@@

### Write excel

@@

### Write excel + sheets

@@

### Write excel - memory issue

@@

@@

@@

```{r}
spo_ces <- ic %>% 
  filter(grepl("Další|Jiná|Lapidárium|Etnografická", zdroj))
```

@@

>>>>

### TXT

read.table = [function](https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/read.table)
read.csv = [function](https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/read.table)
read_csv = [function]()

~ ENCODING ~

fileEncoding="cp1250"
encoding="utf-8"

#### Read txt

```{r}
df <- read.table("D:/VURV/R_hacks/data/simple_table_4read.txt", header = TRUE, sep = ";")
```

#### Read csv

@@

```{r}
df <- read.csv("D:/VURV/R_hacks/data/simple_table_4read.csv", header = TRUE, sep = ";")
```

### XLSX

 ~ PACKAGES ~

 readxl, xlsx, openxlsx

 ~ FUNCTIONS ~

 readxl::read_excel()
 xlsx::read.xlsx()

 xlsx::write.xlsx()

#### Read excel

df <- read_excel("D:/VURV/R_hacks/data/simple_table_4read.xlsx" , sheet = 2, header = TRUE)
```

@@

#### Write excel

@@

#### Write excel + sheets

@@

#### Write excel - memory issue

@@

### Conditional SAVE

f2s <- paste0(path_kk, proj, "_allXLSX.csv")

if(file.exists(f2s)){
  print("You've been here before.")
  rm(f2s)
} else {
  write.table(df_filled, file = f2s, 
            quote = T, row.names = F, 
            sep = ";", dec = ",", 
            na = "", fileEncoding="cp1250")
  rm(f2s)
  print("File saved!")
  }
```

@@

#### Remove multiple colums

typicky nazvy sloupcu z biolibu

```{r}
nam <- colnames(bl_sp)
noquote(nam)
nam2 <- paste0("-", nam)
noquote(nam2)
nam3 <- paste0(nam2, collapse = ",")
nam3
```

@@

spo_ces <- ic %>% 
  filter(grepl("Další|Jiná|Lapidárium|Etnografická", zdroj))
```

@@

## L O O P S

### @@@ Read all files from folder

1) make a list of files

```{r}
Lxls <- list.files(path_data, pattern = "*.xlsx")  # *.txt *.jpg ...
```

2) create expty dataframe and specify col namens

```{r}
df_filled <- data.frame()
# df_filled <- data.frame(matrix(ncol = 15, nrow = 0))

colnam <- c("prirc","rok_prir","dat_nabyt")
```

3) read files and bind them

for(i in Lxls) {
  a <- read_excel(paste0(path_data, i),
                  col_types = "text",
                  range = cell_cols("A:P"), na = " ") %>% # skip = 2 prepsano rangem
    slice(-1)
  names(a) <- colnam
  a$zdroj <- paste(i) # add source txt file name
  # print(count(a))
  df_filled <- rbind(df_filled, a)
  print(i)
  # rm(a)
}
```

### 2023-10-10 _GEN_axi_piskoviste.Rmd
CES MUZEUMDC - reorg.

require(xml2)

@@

>>>>

# packages and path

install.packages("XML")
require(XML)
install.packages("xml2")
require(xml2)
install.packages("methods")
require(methods)
install.packages("tibble")
require(tibble)

path_dmz <- "C:/Users/krizova/Documents/R/piskoviste/dmz/"

x1 <- XML::xmlParse(paste0(path_dmz, "2312116.xml")) # value
x2 <- xml2::read_xml(paste0(path_dmz, "2312116.xml")) # list

library(xml2)
library(tidyverse)

recs <- xml_find_all(x2, "//name")

@@

# NGP

zkusebni nacteni zvlastne formatovaneho CSV

dat <- read.table(paste0(path_snd, "NGP/pk_data.txt"), header = TRUE, sep = ",", fill = T, fileEncoding="cp1250")
nab <- read.csv(paste0(path_snd, "NGP/nabyti.csv"), header = TRUE, sep = ",", fill = T, fileEncoding="cp1250")

# 9.10.2023

dat <- read.table("M:/03 klienti/narodni galerie - NG/ng zdroj dat/2023-SE/CKS komplet/CKS/sb_data.csv",
                  fill = T, sep = ",", header = T)

                header = TRUE, sep = ",", fill = T, quote = "", fileEncoding="utf-8")

x <- dat %>% 
  filter(str_detect(X, "[:alpha:]"))

x <- dat %>% 
  filter()

```{r POSL SMAZAT}
 mutate(
         Mcislo = ifelse(str_detect(cislo, "\\/"), gsub("\\/.*", "", cislo), cislo),
         Mcislo0 = str_pad(Mcislo, 4, pad = "0"),
         Mrok = ifelse(str_detect(cislo, "\\/"), gsub(".*\\/", "", cislo), NA),
         Mrok19 = ifelse(str_detect(Mrok, "^\\d\\d\\s|^\\d\\d$"), paste0("19", Mrok), Mrok),
         MCES_par = paste0(Mcislo0, "/", Mrok19)) %>%
